Operational constraints and closing civic space: International NGO restrictions and China’s Overseas NGO law

### 2023-12-07 manuscript/manuscript.qmd
minor revision in case study section

TNC and Greenpeace are both well-known environmental INGOs. Both are based in Western countries and both have offices around the world. They each began working in China around the turn of the century, with TNC in establishing an office in Beijing in 1998 and Greenpeace in 2001. Despite these similarities, the two INGOs have had quite different experiences to sustain their legal presence in China after the ONGO Law took effect. TNC obtained RO status in May 2017, partnering with the National Forestry and Grassland Administration as their PSU, and receiving authorization to work in 27 provinces. In contrast, Greenpeace has not received RO status and has instead filed 67 temporary activities, filing new documentation nearly ten times each year.

There are several drawbacks to having to constantly file temporary activities [@Corsetti:2019], resulting in far more limited INGO operational space than under RO regulations (recall @fig-ro-ta-pathways). Securing a new partnership with a Chinese organization for each activity and co-filing the requisite paperwork imposes substantial administrative costs. This approach also creates high uncertainty, since there is no guarantee that a current partner organization would agree to repeat their supervisory role for the next project. Moreover, INGO programming has an inherent short-term focus, given that temporary activities cannot last longer than a year. Finally, temporary activities need to focus on safe activities with few political risks in order to attract potential Chinese partners.

>>>>

TNC and Greenpeace are both well-known environmental INGOs. Both are based in Western countries and have offices around the world. They each began working in China around the turn of the century, with TNC in establishing an office in Beijing in 1998 and Greenpeace in 2001. Despite these similarities, the two INGOs have had quite different experiences to sustain their legal presence in China after the ONGO Law took effect. TNC obtained RO status in May 2017, partnering with the National Forestry and Grassland Administration as their PSU, and receiving authorization to work in 27 provinces. In contrast, Greenpeace has not received RO status and has instead filed 67 temporary activities, filing new documentation nearly ten times each year.

Essentially, constantly filing temporary activities has become a circumvent for INGOs like Greenpeace to maintain long-term presence in China [@Corsetti:2019]. However, choosing a legal channel mostly meant for one-off programs results in far more limited INGO operational space than under RO regulations (recall @fig-ro-ta-pathways). Securing a new partnership with a Chinese organization for each activity and co-filing the requisite paperwork imposes substantial administrative costs. This approach also creates high uncertainty, since there is no guarantee that a current partner organization would agree to repeat their role for the next project. Moreover, INGO programming has an inherent short-term focus, given that temporary activities cannot last longer than a year. Finally, temporary activities need to focus on safe activities with few political risks in order to attract potential Chinese partners.
